---
title: "Model org -- general"
author: "Eli"
date: "March 8, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r libraries}
library( here )
library( tidyverse )
library( reshape2 )
library( knitr )
library( UpSetR )
```

```{r set_figure_output}
figure_directory <- paste0( here::here( 'figures' ), '/' )
figure_directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	fig.path = figure_directory,
	fig.keep = 'all',
	dpi = 300,
	fig.width = 11,
	fig.height = 8.5,
	#dev = 'tiff'
	dev = 'png'
)
```

```{r ggplot_modifiers}
source( file = here::here( "ggplot_modifiers.R" ) )
```

```{r helper_tbl}
order_levels <- c( 'Sc', 'Ce', 'Dm', 'Dr', 'Mm', 'Rn', 'Hs' )

samples <- read_tsv( file = here::here( "names_experiments_table.txt" ) ) %>%
  filter( global == TRUE ) %>%
  mutate( .data = ., abbreviation = factor( abbreviation, levels = order_levels ) )

short_name_levels <- samples %>%
  as.data.frame( . ) %>%
  column_to_rownames( df = ., var = 'abbreviation' ) %>%
  .[ order_levels, 'short_name' ]
```

# General sites information

```{r total_sites_table}
total_sites <- tibble()

for (idx in 1:dim( samples )[1]) {
  curr_name <- samples$filename[ idx ]
  curr_fname <- paste0( curr_name, "_site_count.csv" )
  curr_fpath <- here::here( 'R_analysis', curr_fname )
  
  temp <- read_csv( file = curr_fpath ) %>%
    cbind( ., short_name = samples$short_name[ idx ] )
  
  total_sites <- rbind( total_sites, temp )
}

total_sites <- mutate( .data = total_sites, percent_unique = sprintf( fmt = "%.2f", ( unique_sites / total_sites ) * 100.0 ) )

sum_total <- sum( total_sites$total_sites )
sum_unique <- sum( total_sites$unique_sites )
```

## Table - sites identified
```{r table_sites_general}
select( .data = total_sites, short_name, total_sites, unique_sites, percent_unique ) %>%
  kable( . )
```

Total sites identified: `r sum_total`<br>
Unique sites identified: `r sum_unique`<br>
Percent unique: `r sprintf( fmt = "%.2f", (sum_unique / sum_total ) * 100. )`<br>

```{r genome_sizes}
genome_sizes = tibble()

for (idx in 1:dim( samples )[1]) {
  curr_name <- samples$filename[ idx ]
  curr_fname <- paste0( curr_name, "_genome_size.csv" )
  curr_fpath <- here::here( 'R_analysis', curr_fname )
  
  temp <- read_csv( file = curr_fpath ) %>%
    cbind( ., short_name = samples$short_name[ idx ] )
  
  genome_sizes <- rbind( genome_sizes, temp )
}
```

```{r site_per_mb}
sites_per_mb <- merge( x = total_sites, y = genome_sizes, by = "short_name" ) %>%
  mutate( .data = ., total_per_mb = sprintf( fmt = "%.2f", total_sites / MbGenome ) ) %>%
  mutate( .data = ., unique_per_mb = sprintf( fmt = "%.1f", unique_sites / MbGenome ) ) %>%
  arrange( .data = ., MbGenome )
```

## Table - sites per megabase of genome
```{r table_sites_per_megabase}
select( .data = sites_per_mb, -MbGenome ) %>%
  kable( . )
```

Avg sites / Mb: `r mean( as.numeric( sites_per_mb$total_per_mb ) ) %>% sprintf( fmt="%.2f", . )`<br>
Avg unique sites / Mb: `r mean( as.numeric( sites_per_mb$unique_per_mb ) ) %>% sprintf( fmt = "%.1f", . )`<br>

Site every: `r (1E6 / (mean( as.numeric( sites_per_mb$total_per_mb ) )) ) %>% sprintf( fmt="%.1f", . )`<br>
Unique site every: `r (1E6 / (mean( as.numeric( sites_per_mb$unique_per_mb ) )) ) %>% sprintf( fmt="%.1f", . )`<br>

# Gene overlap analysis
```{r gene_overlaps}
geneCuts <- tibble()

for (idx in 1:dim( samples )[1]) {
  curr_name <- samples$filename[ idx ]
  curr_fname <- paste0( curr_name, "_cutsPerGene.csv.gz" )
  curr_fpath <- here::here( 'R_analysis', curr_fname )
  
  temp <- read_csv( file = curr_fpath )
  
  geneCount <- pull( .data = temp, gene_id ) %>%
    unique( . ) %>%
    length( . )
  
  genesCut <- filter( .data = temp, TotalSites > 0 ) %>%
    pull( .data = ., gene_id ) %>%
    unique( . ) %>%
    length( . )
  
  genesUniqueCut <- filter( .data = temp, UniqueSites > 0 ) %>%
    pull( .data = ., gene_id ) %>%
    unique( . ) %>%
    length( . )
  
  fractionCut <- genesCut / geneCount
  
  fractionUniqueCut <- genesUniqueCut / geneCount
  
  medPerGene <- median( temp$TotalSites )
  medUniquePerGene <- median( temp$UniqueSites )
  
  geneCuts <- data.frame( filename = curr_name, 
              Genes = geneCount,
              Cut = fractionCut,
              UniquelyCut = fractionUniqueCut,
              MedSites = medPerGene,
              MedUniqSites = medUniquePerGene ) %>%
    rbind( geneCuts, . )
}
```

## Table -- cuts per gene
```{r cuts_table}
geneCuts %>%
  mutate( .data = ., Cuts = sprintf( "%.2f", Cut * 100.0 ) ) %>%
  mutate( .data = ., UniqueCuts = sprintf( "%.2f", UniquelyCut * 100.0 ) ) %>%
  merge( x = ., y = select( .data = samples, filename, short_name, abbreviation ) ) %>%
  mutate( .data = ., short_name = factor( short_name, levels = short_name_levels ) ) %>%
  select( .data = ., short_name, Genes, Cuts, UniqueCuts, MedSites, MedUniqSites ) %>%
  arrange( short_name ) %>%
  kable( . )
```

```{r range_of_cutters}
low_per_gene <- min( geneCuts$MedSites )
high_per_gene <- max( geneCuts$MedSites )
low_uniq_per_gene <- min( geneCuts$MedUniqSites )
high_uniq_per_gene <- max( geneCuts$MedUniqSites )
```

Standard sites: `r low_per_gene` to `r high_per_gene`<br>
Unique sites: `r low_uniq_per_gene` to `r high_uniq_per_gene`<br>

# PAM site analysis
```{r pams}
pam_sites = tibble()

for (idx in 1:dim( samples )[1]) {
  curr_name <- samples$filename[ idx ]
  curr_fname <- paste0( curr_name, "_pam_count.csv" )
  curr_fpath <- here::here( 'R_analysis', curr_fname )
  
  temp <- read_csv( file = curr_fpath ) %>%
    cbind( ., short_name = samples$short_name[ idx ] )
  
  pam_sites <- rbind( pam_sites, temp )
}

pam_sites <- pam_sites %>%
  rename( .data = ., Counts = AllPam ) %>%
  spread( data = ., key = 'Unique', value = 'Counts' ) %>%
  rename( .data = ., Unique = `TRUE` ) %>%
  rename( .data = ., NonUnique = `FALSE` ) %>%
  mutate( .data = ., Total = Unique + NonUnique )

melted_pams <- melt( pam_sites )

pam_sums <- melted_pams %>%
  group_by( short_name, variable ) %>%
  summarize(
    sums = sum( value )
  )

melted_pams <- melted_pams %>%
  rename( .data = ., count = value ) %>%
  merge( x = ., y = pam_sums, by = c( 'short_name', 'variable' ) ) %>%
  mutate( .data = ., fractional = count / sums )
```

## Figure - PAM usage
```{r stacked_bar_chart}
melted_pams %>%
  filter( variable != "NonUnique" ) %>%
  merge( x = , y = select( samples, abbreviation, short_name ) ) %>%
  select( .data = ., abbreviation, variable, PAM, fractional ) %>%
  ggplot( data = ., mapping = aes( x = abbreviation, y = fractional, fill = PAM )) +
  theme_bw() +
  geom_bar( stat = 'identity', colour = 'black' ) +
  facet_wrap( ~variable ) +
  ylab( "Fraction of PAM sites" ) +
  xlab( "Species" ) +
  gg_bigger_texts +
  scale_fill_manual( values = reds4Palette ) +
  theme( axis.text.x = element_text( face = 'italic' ) ) +
  gg_no_x_grid +
  gg_no_y_grid
```

```{r stacked_bar_chart_kable}
melted_pams %>%
  filter( variable != "NonUnique" ) %>%
  kable( . )
```

# PAM GC content analysis
```{r load_gc_content}
gc_content = tibble()

for (idx in 1:dim( samples )[1]) {
  curr_name <- samples$filename[ idx ]
  curr_fname <- paste0( curr_name, "_gcPerc.txt" )
  curr_fpath <- here::here( 'gc_content', curr_fname )
  
  temp <- read_csv( file = curr_fpath, progress = FALSE, col_names = 'percent_gc' ) %>%
    cbind( ., short_name = samples$short_name[ idx ] )
  
  gc_content <- rbind( gc_content, temp )
}
```

```{r pam_gc_content}
pam_gc <- pam_sites %>%
  filter( .data = ., PAM %in% c( 'TTCC', 'TTCG' ) ) %>%
  group_by( short_name ) %>%
  summarize(
    gc_count = sum( Total ) ) %>%
  merge( x = ., y = filter( .data = pam_sums, variable == "Total" ), by = 'short_name' ) %>%
  select( .data = ., -variable ) %>%
  mutate( .data = ., fraction_gc = gc_count / sums ) %>%
  merge( x = ., y = gc_content, by = 'short_name' )

binom_for_purrr <- function( input ) {
  model = with( input, binom.test( x = gc_count, n = sums, p = percent_gc, alternative = 'less' ) )
  data.frame( P_binom = model$p.value, estimate = model$estimate ) %>%
    return( . )
}

pam_gc_test <- pam_gc %>%
  group_by( short_name ) %>%
  nest( data = . ) %>%
  mutate( .data = ., binom_stuff = map( .x = data, .f = binom_for_purrr ) ) %>%
  unnest( . ) %>%
  arrange( .data = ., short_name )

pam_gc_test$P_adj = p.adjust( pam_gc_test$P_binom )
```

## Table PAM stats
```{r pam_kable}
kable( pam_gc_test )

pam_gc %>%
  mutate( .data = ., pam_usage = case_when(
    fraction_gc < percent_gc ~ "LessThanExpected",
    TRUE ~ "GreaterOrEqualToExpected"
  )) %>%
  arrange( short_name ) %>%
  kable( . )
```

```{r updated_pam_sprintf}
pam_gc_test %>%
  mutate( .data = ., Diff = case_when(
    fraction_gc < percent_gc ~ "LessThanExpected",
    TRUE ~ "GreaterOrEqualToExpected"
  )) %>%
  rename( .data = ., Species = short_name, FractionGCpam = fraction_gc ) %>%
  rename( .data = ., GenomeGC = percent_gc ) %>%
  mutate( .data = ., Pvalue = sprintf( fmt = "%.2e", P_binom ) ) %>%
  mutate( .data = ., FDR_pvalue = sprintf( fmt = "%.2e", P_adj ) ) %>%
  select( .data = ., Species, Diff, GenomeGC, estimate, Pvalue, FDR_pvalue ) %>%
  arrange( Species ) %>%
  kable( . )
```

# Session info
```{r versions}
Sys.time()

getwd()
  
sessionInfo()
```
